---
title: "Cutner-Nina_Intermediate-elective-01"
author: "Nina Cutner"
format: pdf
date: May 10, 2026
execute:
  message: false
  warning: false
---

# Set up

## Packages

```{r}
#| label: packages

# loading packages for data import, cleaning, wrangling, and visualization
library(tidyverse)

# loading package to clean column names
library(janitor)

# loading package for file paths
library(here)

# loading package for working with dates
library(lubridate)

# loading package for arranging multiple plots or plot elements
library(patchwork)
```

## Data

```{r}
#| label: data

# reading in bird monitoring data
birds <- read_csv(
  here("data", "birds copy.csv")
)

# reading in water quality metadata
metadata <- read_csv(
  here("data", "NCOS_YSI_Water_Quality_Monitoring_0 copy.csv")
)

# reading in Venoco Bridge water surface elevation data
venoco_water <- read_csv(
  here("data", "venoco-water copy.csv")
)

```

# 1. Explaining inspiration

The visualization I chose for my inspiration is Steven Ponce's, "Coal falls. Renewables rise." The visualization uses contrasting line trends, direct labeling, and annotations to tell a story about changing energy dominance through time.

This visualization makes sense for my dataset because I am also examining temporal changes across bird species at NCOS. Similar to the inspiration figure, my visualization will use contrasting abundance trends, annotations, and highlighted lines to emphasize how bird presence changes throughout the water year.

My visualization will display time on the x-axis and bird abundance on the y-axis. Different bird species will be represented using contrasting colored lines, while annotations will highlight important seasonal or ecological patterns in the data.



